Improving Emotion Detection through Real-Time Translation of Text to ML Models Trained in Different Languages

### 2023-08-14 paper/appendix/code.tex
Renaming for Submission

\section{Appendix -- Application \& Code Details}
Hello World - Application \& Code Details
\subsection{Application}
\label{appendix:code_application}
Hello World - Application
\clearpage

\subsection{Code Details}
Hello World - Code Details

@@

\clearpage

>>>>

\section{Appendix -- Application \& Code Details}
In this appendix we aim to supply details to the code and application blocks of this project. It should be noted that some of the code has been edited to fit on these pages and might not represent verbatim the code in the GIT repository.

All code was developed and test using a MacBook PRO running \texttt{Ventura 13.4.1}. Since PySpark was developed for cluster processing it may seem counter intuitive to build PySpark code for a Mac; however the intent is that some of this research can be further used on clusters in the future.

\subsection{PySpark Session Setup}
\label{appendix:py-spark-setup}
\begin{minted}[fontsize=\small]{python}
# Setup Spark Session
from pyspark.sql import SparkSession

# Get Imports Needed
from pyspark.sql.functions import col, udf, regexp_replace, lower
from pyspark.sql import functions as F

# Get Datatypes needed for DataFrame manipulation
from pyspark.sql.types import IntegerType, StringType, ArrayType

# Other Imports
import re  # Import the "re" module for regular expressions

# Setup Spark Session
sc = SparkSession \
        .builder \
        .master("local[*]") \
        .appName("Qualyfing-Exam") \
        .getOrCreate()

# Print Spark Version being run
print("Spark V: ", sc.version)
\end{minted}
\clearpage

@@

Flask is a lightweight and simple to use web framework in Python, allowing developers to create web applications and APIs. Due to it's simplicity and components design, Flask allowed us to rapidly build and interactive web service fro demonstration purposes on this project.
\begin{minted}[fontsize=\small]{python}
# Setup API 
from flask import Flask, request, jsonify
from googletrans import Translator

# Setup
from transformers import DistilBertTokenizer
from pyspark.ml.pipeline import PipelineModel

# Initialize the AutoTokenizer
tokenizer = DistilBertTokenizer.from_pretrained("distilbert-base-multilingual-cased")

# Load Saved Models
model_en_rfc = PipelineModel.load("./models/model_en_rfc.model")
model_en_nb = PipelineModel.load("./models/model_en_nb.model")
model_en_lr = PipelineModel.load("./models/model_en_lr.model")
model_de_rfc = PipelineModel.load("./models/model_de_rfc.model")
model_de_nb = PipelineModel.load("./models/model_de_nb.model")
model_de_lr = PipelineModel.load("./models/model_de_lr.model")

app = Flask(__name__)
translator = Translator()

# Setup Emotion Key Reverse
emotions = {v: k for k, v in emotion_key.items()}

@app.route('/predict', methods=['GET'])
def transform_data():
    sentence_en = request.args.get('sentence_en')

    if not sentence_en:
        return jsonify({'error': 'No Sentence (EN) provided!'})

    sentence_de = translator.translate(sentence_en, dest="de").text
    print(f"English: {sentence_en}")
    print(f"German: {sentence_de}")

    words_en = tokenizer.tokenize(sentence_en)
    words_de = tokenizer.tokenize(sentence_de)

    # Create Englsih & German Dataframe
    df_en_api = sc.createDataFrame([(sentence_en, words_en)], ["sentence", "words"])
    df_de_api = sc.createDataFrame([(sentence_de, words_de)], ["sentence", "words"])

    predictions_en_rfc = model_en_rfc.transform(df_en_api)
    predictions_en_nb = model_en_nb.transform(df_en_api)
    predictions_en_lr = model_en_lr.transform(df_en_api)
    predictions_de_rfc = model_de_rfc.transform(df_de_api)
    predictions_de_nb = model_de_nb.transform(df_de_api)
    predictions_de_lr = model_de_lr.transform(df_de_api)

    #Build Response
    resp = {
        "sentence": {
            "english":sentence_en,
            "german":sentence_de,
        },
        "predictions": {
            "en_rfc": emotions.get(predictions_en_rfc.collect()[0]["prediction"]),
            "en_nb":  emotions.get(predictions_en_nb.collect()[0]["prediction"]),
            "en_lr":  emotions.get(predictions_en_lr.collect()[0]["prediction"]),
            "de_rfc": emotions.get(predictions_de_rfc.collect()[0]["prediction"]),
            "de_nb":  emotions.get(predictions_de_nb.collect()[0]["prediction"]),
            "de_lr":  emotions.get(predictions_de_lr.collect()[0]["prediction"]),
        },
    }

    return jsonify(resp)

if __name__ == '__main__':
    app.run(host='127.0.0.1', port=8080)
\end{minted}
\clearpage

### 2023-08-14 paper/appendix/dataset.tex
Renaming for Submission

\subsection{Translation Setup}
All example code was created using Jupyter Notebooks.
\begin{minted}[fontsize=\small]{python}
# *********************
# *** Setup Section ***
# *********************

import pandas as pd

# Emotion DataFrame
# This was predtermined by review of the emotion from English to German
emotion_key = {
    "boredom": "---",
    "love": "Vertrauen",
    "relief": "---",
    "fun": "---",
    "hate": "Ekel",
    "neutral": "Unklar",
    "anger": "Ärger",
    "happiness": "Freude",
    "surprise": "Überraschung",
    "sadness": "Traurigkeit",
    "worry": "Angst",
    "enthusiasm": "Antizipation",
    "empty ": "---",
    "---": "Keine"
}

@@

\subsection{Details on English Dataset}
\label{appendix:dataset_english}
The English dataset is based on tweets\footnote{Tweets from Twitter} that are saved as a \texttt{csv}\footnote{csv = Comma Separated Values} file. The following code was use to filter and process the file for use in this project.
\begin{minted}[fontsize=\small]{python}
# *********************************
# *** English Data Preparations ***
# *********************************

import pandas as pd
from tqdm import tqdm

@@

# Merge German Emotions onto English
df_en = pd.merge(df_en, df_emotions, on='emotion_en', how='left')

# Add German Sentence Column
df_en["sentence_de"] = ""

# Randomly select 1500 rows
df_en = df_en.sample(n=1500, random_state=2023)

@@

\subsection{Details on German Dataset}
\label{appendix:dataset_german}
The German dataset was downloaded from the ETH's\footnote{ETH (German: Eidgenössische Technische Hochschule) or Swiss Federal Institute of Technology} servers via "link" as a \texttt{JSON} file. Since the the file was \texttt{JSON} we will need to filter and explode the file based on teh emotions that are attached to it.
\begin{minted}[fontsize=\small]{python}
# *********************************
# *** German Data Preparations ***
# *********************************

import pandas as pd
from tqdm import tqdm

@@

# Merge Englsih Emotions onto German
df_de = pd.merge(df_de, df_emotions, on='emotion_de', how='left')

# Add English Sentence Column
df_de["sentence_en"] = ""

# Randomly select 1500 rows
df_de = df_de.sample(n=1500, random_state=2023)

>>>>

\subsection{Translation Setup}
All example code was created using Jupyter Notebooks.
\begin{minted}[fontsize=\small]{python}
import pandas as pd

# This was predtermined by review of the emotion from English to German
emotion_key = {
    "boredom": "---",
    "love": "Vertrauen",
    "relief": "---",
    "fun": "---",
    "hate": "Ekel",
    "neutral": "Unklar",
    "anger": "Ärger",
    "happiness": "Freude",
    "surprise": "Überraschung",
    "sadness": "Traurigkeit",
    "worry": "Angst",
    "enthusiasm": "Antizipation",
    "empty ": "---",
    "---": "Keine"
}

@@

\subsection{Details on English Dataset Parsing and Formatting}
\label{appendix:dataset_english}
The English dataset is based on tweets\footnote{Tweets from Twitter} that are saved as a \texttt{csv}\footnote{csv = Comma Separated Values} file. The following code was use to filter and process the file for use in this project.
\begin{minted}[fontsize=\small]{python}
# *********************************
# *** English Data Preparations ***
# *********************************
import pandas as pd
from tqdm import tqdm

@@

# Merge German Emotions onto English
df_en = pd.merge(df_en, df_emotions, on='emotion_en', how='left')
df_en["sentence_de"] = "" # Add German Sentence Column
df_en = df_en.sample(n=1500, random_state=2023) # Randomly select 1500 rows

@@

\subsection{Details on German Dataset Parsing and Formatting}
\label{appendix:dataset_german}
The German dataset was downloaded from the ETH's\footnote{ETH (German: Eidgenössische Technische Hochschule) or Swiss Federal Institute of Technology} servers via "link" as a \texttt{JSON} file. Since the the file was \texttt{JSON} we will need to filter and explode the file based on teh emotions that are attached to it.
\begin{minted}[fontsize=\small]{python}
# *********************************
# *** German Data Preparations ***
# *********************************
import pandas as pd
from tqdm import tqdm

@@

# Merge English Emotions onto German
df_de = pd.merge(df_de, df_emotions, on='emotion_de', how='left')
df_de["sentence_en"] = "" # Add English Sentence Column
df_de = df_de.sample(n=1500, random_state=2023) # Randomly select 1500 rows
